### 2021-10-15 chapter2/chapter.tex
update hhat reference, add me references

Here, the construction of the approximating QBD-RAP is developed intuitively, from a stochastic modelling perspective. The key observation is that, given \(\{\varphi(t)\}\) is constant, then \(\{X(t)\}\) moves deterministically. Upon discretising the state space of the level process into intervals of width \(\Delta\), then, given \(X(t)\), the distribution of time it takes for \(\{X(t)\}\) to leave a given interval on the event that \(\{\varphi(t)\}\) remains in the same phase, is deterministic. We model this deterministic behaviour approximately by concentrated matrix exponential distributions (matrix exponential distributions with low-variance). As in \cite{bo2013}, the discretisation of the state space of \(\{X(t)\}\) corresponds to the level process of a QBD-RAP. Unlike  \cite{bo2013}, however, here we do not take \(\Delta \to 0\). Rather, we suppose that the variance of the concentrated matrix exponential distributions we use to approximate deterministic events gets small. This necessarily means that the order of the matrix exponential increases.

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{chapter2/figs/ME_residual_life_density.pdf}
\caption{The density function for a \emph{concentrated matrix exponential of order 21} from \citep{hhat2020} (blue) and corresponding density functions of the residual lives, \(R_{0.3}\) (red), \(R_{0.6}\) (blue). Observe how the density function of the \(Z_i\) (blue) to approximate a point mass at \(\Delta=1\), while the density functions of \(R_{0.3}\) (red) and \(R_{0.6}\) (blue) approximate point masses at \(0.7\) and \(0.4\), respectively. }
\label{fig:residual distributions}
\end{figure}

@@

\subsection{Computing \(\bs D\)}
In practice we use the class of concentrated matrix exponential distributions (CMEs) found numerically in \citep{hhat2020}. For this class of CMEs, we take the index \(p\) to be the order of the representation. Moreover, we take \(p\) to be odd. The justification for considering representations of odd orders only is that the variance of CME representations of orders \(2p\) and \(2p-1\) are relatively similar and therefore have similar abilities to represent the delta function \citep{hhat2020}. Hence, if we construct a QBD-RAP approximation with a representation of order \(2p\) we expect it to perform similarly to an approximation constructed with representations of order \(2p-1\). However, the computational cost of the latter is lower, so we opt for the order \(2p-1\) representation.

@@

Computationally efficient expressions for \(e^{\bs{S}t}\bs e\) and \(\bs a(t)\) are provided in \citep{hhat2020}.

>>>>

Here, the construction of the approximating QBD-RAP is developed intuitively, from a stochastic modelling perspective. The key observation is that, given \(\{\varphi(t)\}\) is constant, then \(\{X(t)\}\) moves deterministically. Upon discretising the state space of the level process into intervals of width \(\Delta\), then, given \(X(t)\), the distribution of time it takes for \(\{X(t)\}\) to leave a given interval on the event that \(\{\varphi(t)\}\) remains in the same phase, is deterministic. We model this deterministic behaviour approximately by concentrated matrix exponential distributions (matrix exponential distributions with low-variance) \cite{ert2006,hstz2016,ert2006,hht2020,mt2021}. As in \cite{bo2013}, the discretisation of the state space of \(\{X(t)\}\) corresponds to the level process of a QBD-RAP. Unlike  \cite{bo2013}, however, here we do not take \(\Delta \to 0\). Rather, we suppose that the variance of the concentrated matrix exponential distributions we use to approximate deterministic events gets small. This necessarily means that the order of the matrix exponential increases.

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{chapter2/figs/ME_residual_life_density.pdf}
\caption{The density function for a \emph{concentrated matrix exponential of order 21} from \citep{hht2020} (blue) and corresponding density functions of the residual lives, \(R_{0.3}\) (red), \(R_{0.6}\) (blue). Observe how the density function of the \(Z_i\) (blue) to approximate a point mass at \(\Delta=1\), while the density functions of \(R_{0.3}\) (red) and \(R_{0.6}\) (blue) approximate point masses at \(0.7\) and \(0.4\), respectively. }
\label{fig:residual distributions}
\end{figure}

@@

\subsection{Computing \(\bs D\)}
In practice we use the class of concentrated matrix exponential distributions (CMEs) found numerically in \citep{hht2020}. For this class of CMEs, we take the index \(p\) to be the order of the representation. Moreover, we take \(p\) to be odd. The justification for considering representations of odd orders only is that the variance of CME representations of orders \(2p\) and \(2p-1\) are relatively similar and therefore have similar abilities to represent the delta function \citep{hht2020}. Hence, if we construct a QBD-RAP approximation with a representation of order \(2p\) we expect it to perform similarly to an approximation constructed with representations of order \(2p-1\). However, the computational cost of the latter is lower, so we opt for the order \(2p-1\) representation.

@@

Computationally efficient expressions for \(e^{\bs{S}t}\bs e\) and \(\bs a(t)\) are provided in \citep{hht2020}.
